Subtle Gestures with the J!NS MEME

### 2021-08-15 Paper/paper.tex
Write about DataLoggers and ANNs

% TODO
What was our goal?

@@

Since some of the gestures generated barely any responses on the EOG or the IMU readouts
[see figures with an example of a strong and a weak response],
we continued coming up with gestures until we had a total of seven gestures.
The gestures we came up with and their apparent responses can be seen in table \ref{responses}.

\begin{table}
\caption{The quality of the EOG and IMU sensor responses for our gestures.}\label{responses}
\begin{tabular}{|l|l|}
    \hline
    Gesture description & Response quality\\
    \hline
    Pushing glasses up the nose & strong\\
    Readjusting glasses & strong\\
    Thinking pose & weak\\
    Slow nodding & strong\\
    Tilting head sideways & strong\\
    Stroking mustache & weak\\
    Biting lip & weak\\
    Tapping on the nose & strong\\
    Rubbing the nose & strong\\
    Pushing cheek up with a finger & strong\\
    Tapping the glasses & strong\\\hline
\end{tabular}
\end{table}

\subsection{Tools for Recording and Labelling}
% is this something for the implementation section? ~FB
How does the J!ns Meme work and how do we use it?

What data do we collect, how and why?
Itchy Nose tool for automating personalization \cite{10.1145/3174910.3174953}.

\subsection{Neural Network Architecture and Training}
% again - how much of this belongs to implementation? ~FB

>>>>

@@

Since some of the gestures generated barely any responses on the EOG or the IMU readouts
[see figures with an example of a strong and a weak response],
we continued coming up with gestures until we had a total of seven (eleven, if left and
right variants are differentiated) presumably recognizable gestures.
The gestures we came up with and their apparent responses can be seen in table
\ref{tab:responses}.

\begin{table}
\caption{The quality of the EOG and IMU sensor responses for our gestures. Gestures 
        with (L/R) at the end have both a left and a right variant.}\label{tab:responses}
\begin{tabular}{|l|l|}
    \hline
    Gesture description & Response quality\\
    \hline
    Pushing glasses up the nose & strong\\
    Readjusting glasses (L/R)& strong\\
    Thinking pose & weak\\
    Slow nodding & strong\\
    Tilting head sideways (L/R)& strong\\
    Stroking mustache & weak\\
    Biting lip & weak\\
    Tapping on the nose (L/R)& strong\\
    Rubbing the nose & strong\\
    Pushing cheek up with a finger (L/R)& strong\\
    Tapping the glasses (L/R)& strong\\\hline
\end{tabular}
\end{table}

\subsection{Tools for Recording and Labelling}
% TODO include the links via BibTeX
To use the J!ns Meme, we first had to interface with it. The interface would have to allow
both recording and streaming the data from the smartglasses, perferrably.
The J!ns Meme's sensor data can be read either via an iOS, Android or macOS app
or via a Windows app, with the latter requiring a custom bluetooth dongle. For
initial gesture evaluation we used the Android app, but since we set out to have
live classification with the potential to control your computer, and only Windows
machines were available, we chose the Windows app J!ns Meme DataLogger
[https://github.com/jins-meme/ES\_R-DataLogger-for-Windows]. A development kit was
available for Android, macOS and Windows as well
[https://github.com/jins-meme/ES\_R-Development-Kit], but it did not eliminate the
need for the bluetooth dongle on Windows and only implied a bigger overhead for us,
since the DataLogger can both record and forward the J!ns Meme's sensor data to a
file and a TCP port, respectively.

%TODO check correctness, this is from memory alone
The J!ns Meme DataLogger supplies eleven values per cycle, three of which are the
accelerometers' data and the remaining eight being processed EOG data.
Notably, we observed that we had different values available via the Android app
[https://github.com/jins-meme/ES_R-DataLogger-for-Android]. Namely
there were gyroscope values, albeit they were constantly zero, and less EOG data fields.
We found this behavior to be the same on all J!ns Meme smartglasses and Android phones
we had available, but are not certain if it is intended behavior.

Since we wanted our neural network to have the most possible data, we simply included all
eleven values that the DataLogger on Windows provided into our data collection.

The next step after the data collection would have been the labelling of the recorded data.
But with our goal being at least hundreds of samples, manual labelling seemed impractical.
Therefore, we developed a tool that would firstly present our test subjects the
gestures they would have to perform and secondly ask them to press a button to
automatically document the start and end of their gesture. This allowed us to split the
file that the DataLogger would create into multiple files, taking the included time
stamps as a reference.

\subsection{Neural Network Architecture and Training}
For the classification task of telling the gestures apart from each other, we decided on
a long short-term memory (LSTM) neural network. LSTM neural networks, contrary to convolutional neural networks
or multilayer perceptrons, allow for the input data to have different lengths. This
implies that we do not need to add padding to our data samples to bring them all to the
same length. Also, the data we use as input for the network is sequential and a time
series. Recurrent neural networks such as LSTMs can remember what data they were supplied
in the last time steps by having backward edges in their hidden layers, making them
usually a good fit for sequential data.
